Mega-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-08-31 paper/1_Initial/manuscript.tex
manuscript pdf

\documentclass[
  man,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

@@

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

@@

\usepackage[nolongtablepatch]{lineno}
\linenumbers

@@

\title{Meta-analysis of the Interoceptive Accuracy Scale (IAS) Structure
and its Subjective Correlates}

@@

\author{Ana Neves}

\affiliation{
{School of Psychology, University of Sussex}}

\leftheader{Neves}

\abstract{Blabla the abstract blabla.}

@@

\authornote{\par{\addORCIDlink{Ana Neves}{0009-0006-0020-7599}}

\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colback=white, breakable, opacityback=0, leftrule=.75mm, arc=.35mm, bottomrule=.15mm, toprule=.15mm, left=2mm, colframe=quarto-callout-note-color-frame]

@@

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}  Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Ana
Neves:   Project administration, Data curation, Formal
Analysis, Investigation, Visualization, Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed to }
}

\makeatletter
\let\endoldlt\endlongtable
\def\endlongtable{
\hline
\endoldlt
}
\makeatother

@@

\usepackage{lscape}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\resetlinenumber[1]

Interoception is referred to the process of sensing, interpreting and
integrating information pertaining to internal organs, such as the
heart, the lungs or the gut (\citeproc{ref-khalsa2018}{Khalsa, Adolphs,
Cameron, Critchley, Davenport, Feinstein, Feusner, Garfinkel, Lane,
Mehling, Meuret, et al., 2018}). While recent research emphasizes a key
role of interoception in a variety of processes (e.g., emotion
regulation, decision making) and of outcomes (physical and psychological
well being), the field remains clouded by concerns about how
interoception is assessed.

Various measures of interoception have been developed (see
Figure~\ref{fig-measures}), forming a combination of ``objective'' and
``subjective'' assessments (i.e., physiological tasks such as the heart
beat counting or tracking vs.~questionnaires and subjective scales
involving a metacognitive reflection), ``explicit'' and ``implicit''
paradigms (i.e., directing participants' awareness and attention to
interoceptive processes \emph{vs.} measuring interoception unbeknownst
to them), various interoceptive modalities (e.g., cardioception,
respiroception, gastroception) and theoretical dimensions (e.g.,
accuracy, sensitivity, awareness). While there is no consensus as to
which particular approach provides the most accurate and ``pure''
measure of interoception and interoceptive abilities (assuming it is a
unidimensional construct), it is instead plausible that each measure has
strengths and limitations, and a utility dependent on the context and
goal at hand (\citeproc{ref-jahedi2014}{Jahedi \& Méndez, 2014}).

@@

{\caption{{Different ways in which interoception can be
measured.}{\label{fig-measures}}}}

\includegraphics[width=6.4in,height=\textheight]{figures/MeasuringInteroception.jpg}

@@

The use of subjective self-report questionnaires to measure deeply
embodied functions might seem paradoxical. However, recent redefinitions
of interoception emphasize the role of high-level and metacognitive
elaboration of interoceptive information. These redefinitions provide
theoretical grounding to support the idea that some facets of
interoception, including participants' metacognitive beliefs, can be
assessed subjectively (\citeproc{ref-khalsa2018interoception}{Khalsa,
Adolphs, Cameron, Critchley, Davenport, Feinstein, Feusner, Garfinkel,
Lane, Mehling, \& others, 2018};
\citeproc{ref-suksasilp2022towards}{Suksasilp \& Garfinkel, 2022}). This
approach offers useful and interesting measures
(\citeproc{ref-lin2023}{Lin et al., 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019}).

The notion that self-reports might not reflect the same processes as
other interoception measures is important to contextualize the apparent
lack of convergence between measures in the field
(\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022}). While few
studies have explored the correlation between objective interoceptive
measures, such as the Heartbeat Counting Task (HCT,
\citeproc{ref-schandry1981heart}{Schandry, 1981}) and the Heartbeat
Detection Task (HDT, \citeproc{ref-kleckner2015methodological}{Kleckner
et al., 2015}), and correspondent self-report measures, existing
findings typically show weak or no correlations
(\citeproc{ref-brand2023}{Brand et al., 2023}; e.g., task-based accuracy
vs.~self-reported accuracy, \citeproc{ref-murphy2019}{Murphy et al.,
2019}).

Even among self-report measures low correlations have been observed,
suggesting that these measures, while group in the same umbrella term of
self-report assess different facets of interoception. A recent
systematic review by Desmedt et al.
(\citeproc{ref-desmedt2022measures}{2022}) examined whether various
questionnaires designed to assess ``interoceptive sensibility'' truly
measure the same construct. Notably, Garfinkel et al.
(\citeproc{ref-garfinkel2015knowing}{2015}) defines interoceptive
sensibility as the self-reported tendency to focus on and detect
internal sensations, whereas Khalsa, Adolphs, Cameron, Critchley,
Davenport, Feinstein, Feusner, Garfinkel, Lane, Mehling, and others
(\citeproc{ref-khalsa2018interoception}{2018}) defines it more narrowly,
excluding the ability to detect these signals. According to Desmedt et
al. (\citeproc{ref-desmedt2022measures}{2022}), most authors adopt the
latter definition. The review found that these questionnaires measure
related but distinct constructs, leading researchers to mistakenly
treating them as equivalent measures of interoceptive sensibility. A
better understanding of what is being measured with different
questionnaires and dimensions, as well as their potential overlaps with
other constructs (e.g., alexithymia, body awareness), is thus needed to
clarify the role of self-reports in the assessment of interoception.

A recently developed scale with a rapidly growing popularity is the
Interoceptive Accuracy Scale (IAS, \citeproc{ref-murphy2019}{Murphy et
al., 2019}). The IAS consists of 21 Likert-scale items that query how
accurately one can perceive different bodily signals, with one item per
physiological modality such as respiration (\emph{``I can always
accurately perceive when I am breathing fast''}), heart (e.g \emph{``I
can always accurately perceive when my heart is beating fast''}), skin
(e.g \emph{``I can always accurately perceive when something is going to
be ticklish''}), arousal or bodily functions like coughing (e.g
\emph{``I can always accurately perceive when I am going to cough''}) or
urinating (e.g.~\emph{``I can always accurately perceive when I need to
urinate''}). Interestingly, the IAS' statements are about specific
interoceptive behaviours, which is a notable difference with other
popular interoception questionnaires, such as the Multidimensional
Assessment of Interoceptive Awareness scale (MAIA,
\citeproc{ref-mehling2012}{Mehling et al., 2012}; MAIA-2,
\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}), which
contains more general and metacognitive items (e.g., \emph{``I trust my
body sensations''}, \emph{``I can notice an unpleasant body sensation
without worrying about it''}).

Although the original validation study suggested a two-factor structure
for the IAS, the authors underline its acceptable but imperfect fit
(\citeproc{ref-murphy2019}{Murphy et al., 2019, p. 127}), calling on
further investigation of the scale's factor structure. Notably, the
authors did not define these factors, as no clear explanations were
evident. They suggested that the first factor reflects the perception of
interoceptive signals, while the second pertains to signals that may be
difficult to perceive solely through interoceptive information.

\textbf{note to dom}:Morin et al. (\citeproc{ref-morin2016}{2015}) is a
methodological paper on ESEM

Other follow-up studies using confirmatory factor analysis (CFA) and
structural modeling have identified different optimal solutions. Some
studies, like Brand et al. (\citeproc{ref-brand2023}{2023}), reported a
1-factor solution, while Lin et al. (\citeproc{ref-lin2023}{2023}) and
Campos et al. (\citeproc{ref-campos2021}{2021}) found bifactor solutions
- one general factor and a set of lower-level factors
(\citeproc{ref-rodriguez2016evaluating}{Rodriguez et al., 2016}) - to be
the best fit. Notably, the only other study to report a 2-factor
solution was conducted by Koike and Nomura
(\citeproc{ref-koike2023}{2023}), who performed an Exploratory Factor
Analysis (EFA) assuming 2 factors to align with the findings from the
original validation paper.

Discussions have also been focused on specific items. For instance,
Murphy et al. (\citeproc{ref-murphy2019}{2019}) notes that some items
might measure direct interoceptive signals such as cardioception, while
others might capture phenomena not perceivable through interoceptive
signals alone (e.g., ``bruising''; p.~119). Lin et al.
(\citeproc{ref-lin2023}{2023}) also highlights their correlation
analysis, showing five locally dependent pairs and three items (touch,
blood sugar, bruise) with exceptionally high difficulty and low
discrimination. Additionally, Campos et al.
(\citeproc{ref-campos2021}{2021}) reported ``tickle'' to be the only
item that reflected more specific factors than the general factor.
Interestingly, Lin et al. (\citeproc{ref-lin2023}{2023}) reported that
all items of the IAS grouped together using a new approach, Exploratory
Graph Analysis (EGA, \citeproc{ref-golino2017exploratory}{Golino \&
Epskamp, 2017}), to assess convergent and discriminant validity,
providing further evidence for unidimensionality.

The IAS has naturally been compared to other interoception-related
scales, and shows a positive correlations with most facets of the MAIA
(\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}),
except for the Non-Distracting and Not-Worrying subscales
(\citeproc{ref-brand2023}{Brand et al., 2023}). Interestingly, findings
on the correlation between the IAS and the body awareness dimension of
the Body Perception Questionnaire (BPQ-A,
\citeproc{ref-porges1993body}{Porges, 1993}) have been mixed: some
studies report small positive correlations
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023}), while others find
small negative correlations (\citeproc{ref-lin2023}{Lin et al., 2023})
or no correlation at all (\citeproc{ref-murphy2019}{Murphy et al.,
2019}). Small positive correlations have also been observed with the
``observation'' and ``description'' subscales of the Five Facet
Mindfulness Questionnaire (FFMQ, \citeproc{ref-baer2006using}{Baer et
al., 2006}; \citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023}), as well as with the
``non-reactivity'' and ``acting with awareness'' subscales
(\citeproc{ref-koike2023}{Koike \& Nomura, 2023}). Additionally, the IAS
has shown a positive correlation with the interoceptive awareness
subscale of the Eating Disorder Inventory (\citeproc{ref-lin2023}{Lin et
al., 2023}) and a negative correlation with the Interoceptive Confusion
Questionnaire (\citeproc{ref-brand2023}{Brand et al., 2023}; ICQ,
\citeproc{ref-brewer2016alexithymia}{Brewer et al., 2016};
\citeproc{ref-murphy2019}{Murphy et al., 2019}). Lastly, small positive
correlations have also been reported with the Interoceptive Attention
Scale (\citeproc{ref-koike2023}{Koike \& Nomura, 2023}; IATS,
\citeproc{ref-lin2023}{Lin et al., 2023}), though studies have also
found no correlation between these measures
(\citeproc{ref-gabriele2022dissociations}{Gabriele et al., 2022}).
Interestingly, the IAS and the IATS supposedly measure two different
interoceptive processes (i.e., accuracy and attention, respectively)
which contradict Murphy et al. (\citeproc{ref-murphy2019}{2019})
proposed 2x2 framework.

While assessing the validity of an interoception scale can be conceived
as theoretically challenging, several measures have been used to assess
convergent validity for the the IAS, including expected negative
associations with alexithymia (\citeproc{ref-brand2023}{Brand et al.,
2023}; \citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-lin2023}{Lin et al., 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019}), somatic symptoms
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-lin2023}{Lin et al., 2023}), depressive symptoms
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-lin2023}{Lin et al., 2023}), anxiety
(\citeproc{ref-brand2023}{Brand et al., 2023}), neuroticism
(\citeproc{ref-brand2023}{Brand et al., 2023}) and self-esteem
(\citeproc{ref-murphy2019}{Murphy et al., 2019}).

The current study aims at 1) clarifying the structure of the IAS with a
meta-analytic approach that leverages existing data and contrast the
traditional CFA/SEM factor-based analyses with network-based ones such
as EGA; 2) the second part will provide an overview of the dispositional
correlates of the IAS, clarifying the pattern of associations which is
key to better understand the nature, place and role of interoception
questionnaires within a larger context.

>>>>

\documentclass[
  jou,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

@@

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\usepackage{subcaption}

\renewcommand\thesubfigure{\Alph{subfigure}}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

@@

@@

\title{Mega-analysis of the Interoceptive Accuracy Scale (IAS) Structure
and its Subjective Correlates}

@@

\authorsnames[{1},{2},{1},{1},{1,3},{1,4}]{Ana Neves,Giulia
Poerio,Magdalena Pfaff,Robyn Scharte,Raquel Nogueira Arjona,Dominique
Makowski}

\authorsaffiliations{
{School of Psychology, University of Sussex},{},{Sussex Addictions
Research and Intervention Centre, University of Sussex},{Sussex Centre
for Consciousness Science, University of Sussex}}

\leftheader{Neves, Poerio, Pfaff, Scharte, Arjona and Makowski}

\abstract{Blabla the abstract blabla. }

@@

\authornote{\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poerio}{0000-0002-2343-5109}}\par{\addORCIDlink{Magdalena
Pfaff}{0009-0006-2386-7936}}\par{\addORCIDlink{Robyn
Scharte}{blabla}}\par{\addORCIDlink{Raquel Nogueira
Arjona}{0000-0002-5998-701X}}\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}

\par{     \begin{tcolorbox}[enhanced jigsaw, left=2mm, bottomrule=.15mm, breakable, leftrule=.75mm, opacityback=0, colframe=quarto-callout-note-color-frame, arc=.35mm, rightrule=.15mm, toprule=.15mm, colback=white]

@@

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}  Author roles were classified using the Contributor Role
Taxonomy (CRediT; https://credit.niso.org/) as follows:  Ana
Neves:   Data curation, Formal
Analysis, Investigation, Visualization, Writing -- original
draft, Writing -- review \& editing; Giulia Poerio:   bla bla; Magdalena
Pfaff:   Data curation, Writing -- original draft; Robyn Scharte:   bla
bla; Raquel Nogueira Arjona:   bla bla; Dominique Makowski:   Project
administration, Data curation, Formal
Analysis, Investigation, Visualization, Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed
to Dominique
Makowski, Email: \href{mailto:D.Makowski@sussex.ac.uk}{D.Makowski@sussex.ac.uk}}
}

\usepackage{pbalance}
% \usepackage{float}
\makeatletter
\let\oldtpt\ThreePartTable
\let\endoldtpt\endThreePartTable
\def\ThreePartTable{\@ifnextchar[\ThreePartTable@i \ThreePartTable@ii}
\def\ThreePartTable@i[#1]{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.485\textwidth}
\oldtpt[#1]
}
\def\ThreePartTable@ii{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.48\textwidth}
\oldtpt
}
\def\endThreePartTable{
\endoldtpt
\end{minipage}
\twocolumn
\end{figure}}
\makeatother

\makeatletter
\let\endoldlt\endlongtable		
\def\endlongtable{
\hline
\endoldlt}
\makeatother

\newenvironment{twocolumntable}% environment name
{% begin code
\begin{table*}[!htbp]%
\onecolumn%
}%
{%
\twocolumn%
\end{table*}%
}% end code

@@

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{lscape}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\section{Introduction}\label{introduction}

Interoception is referred to the process of sensing, interpreting and
integrating information pertaining to internal organs, such as the
heart, the lungs or the gut
(\citeproc{ref-khalsa2018interoception}{Khalsa et al., 2018}). While
recent research emphasizes a key role of interoception in a variety of
processes (e.g., emotion regulation, decision making) and of outcomes
(physical and psychological well being), the field remains clouded by
concerns about how interoception is assessed.

\subsection{The Interoceptive Assessment
Puzzle}\label{the-interoceptive-assessment-puzzle}

Various measures of interoception have been developed (see
Figure~\ref{fig-measures}), forming a combination of ``objective'' and
``subjective'' assessments (i.e., physiological tasks such as the heart
beat counting or tracking \emph{vs.} questionnaires and subjective
scales involving metacognitive judgments), ``explicit'' and ``implicit''
paradigms (i.e., directing participants' awareness and attention to
interoceptive processes \emph{vs.} measuring interoception unbeknownst
to them), various interoceptive modalities (e.g., cardioception,
respiroception, gastroception) and theoretical dimensions (e.g.,
accuracy, sensitivity, awareness). While there is no consensus as to
which particular approach provides the most accurate and ``pure''
measure of interoception and interoceptive abilities (assuming it is a
unidimensional construct), it is instead plausible that each measure has
strengths and limitations, and a utility dependent on the context and
goal at hand (\citeproc{ref-Desmedt2023accuracy}{Desmedt et al., 2023};
\citeproc{ref-jahedi2014}{Jahedi \& Méndez, 2014}).

@@

{\caption{{The Interoceptive Assessment Puzzle. The different modalities
of interoception (e.g., cardioception) can be assessed directly or
indirectly. Direct assessments can further be subjective or objective,
depending on whether they involve conscious metacognitive appraisals or
more performance-based indices. Interoceptive tasks can be explicit (the
participant is aware of the interoceptive nature of the task and must
consciously attend to interoceptive signals; e.g., the heartbeat
counting task) or implicit (measurements of interoception done
unbeknowst to the participant; e.g., heartbeat evoked potentials
measured during resting state). Indirect assessments evaluate constructs
typically related (and ideally dependent on) to interoceptive processes
or ability (or its deficit).}{\label{fig-measures}}}}

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/MeasuringInteroception.png}

@@

Although the use of subjective self-report questionnaires to measure
deeply embodied functions might seem paradoxical at first, recent
redefinitions of interoception emphasize the role of high-level and
metacognitive elaboration of interoceptive information. These
redefinitions provide theoretical grounding to support the idea that
some facets of interoception, including participants' metacognitive
beliefs, can be assessed subjectively
(\citeproc{ref-khalsa2018interoception}{Khalsa et al., 2018};
\citeproc{ref-suksasilp2022towards}{Suksasilp \& Garfinkel, 2022}).
Moreover, the notion that self-reports might not reflect the same
processes as other interoception tasks might be important to
contextualize the apparent lack of convergence between measures in the
field (\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022}). For
instance, existing findings typically show weak or no correlations
between questionnaires and objective measures, such as the Heartbeat
Counting Task (HCT, \citeproc{ref-schandry1981heart}{Schandry, 1981})
and the Heartbeat Detection Task (HDT,
\citeproc{ref-kleckner2015methodological}{Kleckner et al., 2015}),
including for measures of the same theoretical dimensions
(\citeproc{ref-arslanova2022}{Arslanova et al., 2022};
\citeproc{ref-brand2023}{Brand et al., 2023}; e.g., task-based accuracy
vs.~self-reported accuracy, \citeproc{ref-murphy2019}{Murphy et al.,
2019}). Additionally, even various objective measures assessing in
theory the same interoceptive dimension, such as accuracy, either show
no or weak correlation (respectively, \citeproc{ref-brand2023}{Brand et
al., 2023}; \citeproc{ref-hickman2020relationship}{Hickman et al.,
2020}). Perhaps more surprisingly, low correlations have been observed
even among questionnaires, suggesting (in parallel to major validity
concerns) the potential targeting of different facets related to
interoception.

One striking example concerns the assessment of interoceptive
sensibility, which is broadly defined as the self-reported tendency to
focus on and detect internal sensations
(\citeproc{ref-garfinkel2015knowing}{Garfinkel et al., 2015}), but more
narrowly as the subjective tendency to focus on interoceptive signals,
without necessarily implying detection ability
(\citeproc{ref-khalsa2018interoception}{Khalsa et al., 2018}). A recent
systematic review suggested that various questionnaires designed to
assess interoceptive sensibility may, in fact, measure distinct
constructs, with the risk of researchers treating them as equivalent
despite overall low convergence
(\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022}). Notably,
this review adopted a broad definition of sensibility, incorporating
both interoceptive sensibility and interoceptive self-report scales,
following the eight-facet model by Khalsa et al.
(\citeproc{ref-khalsa2018interoception}{2018}). Several widely used
questionnaires were included in the review, such as the Multidimensional
Assessment of Interoceptive Awareness (MAIA,
\citeproc{ref-mehling2012}{Mehling et al., 2012}; MAIA-2,
\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}), the
Body Perception Questionnaire (BPQ,
\citeproc{ref-porges1993body}{Porges, 1993}), the Private subscale of
the Body Consciousness Questionnaire (PBCS,
\citeproc{ref-miller1981consciousness}{Miller et al., 1981}), the Body
Awareness Questionnaire (BAQ, \citeproc{ref-shields1989body}{Shields et
al., 1989}), and the Eating Disorder Inventory
(\citeproc{ref-garner1983}{Garner et al., 1983}; EDI,
\citeproc{ref-garner1991eating}{Garner, 1991}). The lack of correlations
to moderate correlations among these questionnaires highlight the need
for greater conceptual clarity regarding what each measure captures, how
they relate to different dimensions of interoception, and their
potential overlaps with other constructs, such as alexithymia and body
awareness.

\subsection{The Interoceptive Accuracy Scale
(IAS)}\label{the-interoceptive-accuracy-scale-ias}

Focusing on another dimension of interoception, a recently developed
scale with a rapidly growing popularity is the Interoceptive Accuracy
Scale (IAS, \citeproc{ref-murphy2019}{Murphy et al., 2019}). The IAS
consists of 21 Likert-scale items that query how accurately one can
perceive different bodily signals, with one item per physiological
modality such as respiration (\emph{``I can always accurately perceive
when I am breathing fast''}), heart (\emph{``I can always accurately
perceive when my heart is beating fast''}), skin (\emph{``I can always
accurately perceive when something is going to be ticklish''}), arousal
or bodily functions like coughing (\emph{``I can always accurately
perceive when I am going to cough''}) or urinating (\emph{``I can always
accurately perceive when I need to urinate''}). Appealingly, the IAS'
statements are about specific interoceptive behaviours, which is a
distinct difference with other popular interoception questionnaires,
such as the MAIA-2, which contains more general and metacognitive items
(e.g., \emph{``I trust my body sensations''}, \emph{``I can notice an
unpleasant body sensation without worrying about it''}), as well as
dimensions related to attention regulation (e.g., Not-distracting) or
emotion regulation (e.g., Not-worrying).

{[}\textbf{NOTES}: todd 2022 found no correlation for MAIA and IAS
except with the non-worrying facet + Gaggero found positive correlations
with IAS and MAIA (general){]}

The original validation study suggested a two-factor structure for the
IAS, one reflecting the perception of general interoceptive signals
(urinate, hungry, defecate, thirsty, pain, heart, taste, breathing,
temperature, muscles, affective touch, vomit, sexual arousal), and other
relating to signals that may be difficult to perceive solely through
interoceptive information (itch, tickle, cough, burp, bruise, blood
sugar, sneeze, wind). The authors however underlined its acceptable but
imperfect fit (\citeproc{ref-murphy2019}{Murphy et al., 2019, p. 127}),
and several follow-up studies have indeed identified different optimal
solutions. For instance, Brand et al. (\citeproc{ref-brand2023}{2023})
reported a 1-factor solution, while Lin et al.
(\citeproc{ref-lin2023}{2023}) - using Exploratory Graph Analysis (EGA,
\citeproc{ref-golino2017exploratory}{H. F. Golino \& Epskamp, 2017}) -
and Campos et al. (\citeproc{ref-campos2021}{2021}) found bifactor
solutions (i.e., one general factor above a set of lower-level factors,
\citeproc{ref-rodriguez2016evaluating}{Rodriguez et al., 2016}) to be
the best fit. Using a 2-factors Exploratory Factor Analysis (EFA), Koike
and Nomura (\citeproc{ref-koike2023}{2023}) suggested that the items
could be grouped into cutaneous (itching, tickling, coughing, burping,
affective touch, bruising, passing gas, sneezing, muscle sensations,
sexual arousal, and taste) and visceral sensations (urination,
defecation, hunger, thirst, pain, breathing, fatigue/blood sugar,
temperature, vomiting, and heartbeat).

Discussions have also been focused on specific items. For instance,
Murphy et al. (\citeproc{ref-murphy2019}{2019}) notes that some items
might measure direct interoceptive signals such as cardioception, while
others might capture phenomena not perceivable through interoceptive
signals alone (e.g., ``bruising''; p.~119). Lin et al.
(\citeproc{ref-lin2023}{2023}) additionally highlights five locally
dependent pairs and three items (touch, blood sugar, bruise) with
exceptionally high difficulty and low discrimination, and Campos et al.
(\citeproc{ref-campos2021}{2021}) reported ``tickle'' to be the only
item that reflected more specific factors than the general factor.
Furthermore, localization issues also arose, with both ``itch'' and
``tickle'' corresponding to the same Chinese character, leading to their
collapse into a single item (\citeproc{ref-lin2023}{Lin et al., 2023}).

Regarding its validity, the IAS has naturally been compared to other
interoception-related measures, and shows a positive correlations with
most facets of the MAIA
(\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}),
except for the Not-Distracting and Not-Worrying subscales
(\citeproc{ref-brand2023}{Brand et al., 2023}) - which were previously
highlighted as related to non-interoceptive abilities
(\citeproc{ref-ferentzi2021examining}{Ferentzi et al., 2021}).
Interestingly, findings on the correlation between the IAS and the body
awareness dimension of the BPQ (i.e., BPQ-A) have been mixed: some
studies report small positive correlations
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023}), while others find
small negative correlations (\citeproc{ref-lin2023}{Lin et al., 2023})
or no correlation at all (\citeproc{ref-murphy2019}{Murphy et al.,
2019}). Small positive correlations have also been observed with the
``observation'' and ``description'' subscales of the Five Facet
Mindfulness Questionnaire (FFMQ, \citeproc{ref-baer2006using}{Baer et
al., 2006}; \citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023}), as well as with the
``non-reactivity'' and ``acting with awareness'' subscales
(\citeproc{ref-koike2023}{Koike \& Nomura, 2023}). Additionally, the IAS
has shown a positive correlation with the interoceptive awareness
subscale of the Eating Disorder Inventory (EDI-IA,
\citeproc{ref-lin2023}{Lin et al., 2023}) and a negative correlation
with the Interoceptive Confusion Questionnaire (ICQ,
\citeproc{ref-brewer2016alexithymia}{Brewer et al., 2016}), as reported
by Brand et al. (\citeproc{ref-brand2023}{2023}) and Murphy et al.
(\citeproc{ref-murphy2019}{2019}). Lastly, the correlation with the
Interoceptive Attention Scale (IATS,
\citeproc{ref-gabriele2022dissociations}{Gabriele et al., 2022}) appears
rather small (\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-lin2023}{Lin et al., 2023}).
